---
title: "Metro Phoenix Wildlife Study - Community Dynamics - Model Input Data Preparation"
author: "Jeffrey Haight"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
gc()
```

 
```{r packages, include = FALSE}
library(abind)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(beepr)


```

# Import Datasets
```{r import detection data}
# Three seasons of data from the Metro Phoenix Wildlife Study
    y <- readRDS("~/GitHub/metrophxwildlife-communitydynamics/data/model_inputs/y_s22_j50_occ16day_3season.rds")

# the basic sampling stats
    n.site <- dim(y)[1]   # number of sites
    n.survey <- dim(y)[2]  # number of surveys
    n.season <- dim(y)[3]  # number of seasons
    n.spp <- dim(y)[4]     # number of species
    
    
# Site survey effort
    K <- read.csv("~/GitHub/metrophxwildlife-communitydynamics/data/model_inputs/K_tot_j50_3season.csv", header = T)
    K$effort <- K$season1 +K$season2 + K$season3
  

# Put the data into a collapsed format, with number of by-survey detections
  # this will make it easier to calculate likelihood and reduce the dimensions of the input data (though it creates a new input)
  # number of full surveys of each camera in each seasons
  # same number of surveys per species, so just select the first species
  #K.survey <- apply(y, c(1,3,4), function(x) length(which(!is.na(x))))[,,1]  
  K.survey <- apply(y[,,,1], c(1,3), function(x) length(which(!is.na(x))))
  
  ysum <- apply(y, c(1,3,4), function(x) sum(x, na.rm = TRUE))
  #ysum[,2,]
```

```{r import site covariate data}

# Site covariate data
data.site <- read.csv("~/GitHub/metrophxwildlife-communitydynamics/data/model_inputs/covariates_50camsite.csv") %>% arrange(site)


  # add survey effort columns
  #data.site$K_days <- c(K_wd, K_ww, K_cw)    # number of surveyed days
  
  #data.site$effort <- K_days/max(K_days)     
  
  #data.site <- data.site %>%
  #  group_by(season) %>%
  #  mutate(
  #    effort = K_days/max(K_days)        # proportion of days in each season that were surveyed
  #  )
  
  # add quadratic urbanization terms
  data.site$impquad100m  <- ((data.site$imp100m - mean(data.site$imp100m))/sd(data.site$imp100m))^2
  data.site$impquad1km  <- ((data.site$imp1km - mean(data.site$imp1km))/sd(data.site$imp1km))^2
  data.site$impquad5km  <- ((data.site$imp5km - mean(data.site$imp5km))/sd(data.site$imp5km))^2

  # standardize non-seasonal covariates (e.g. imperviousness)
  data.site <- data.site %>%
          #group_by(season) %>%
          mutate(
            imp100m_std = (imp100m - mean(imp100m))/sd(imp100m),
            imp1km_std = (imp1km - mean(imp1km))/sd(imp1km),
            imp5km_std = (imp5km - mean(imp5km))/sd(imp5km),
            sdhi100m_std = (sdhi100m - mean(sdhi100m))/sd(sdhi100m),
            sdhi1km_std = (sdhi1km - mean(sdhi1km))/sd(sdhi1km),
            sdhi5km_std = (sdhi5km - mean(sdhi5km))/sd(sdhi5km),
            pd100m_std = (pd100m - mean(pd100m))/sd(pd100m),
            pd1km_std = (pd1km - mean(pd1km))/sd(pd1km),
            pd5km_std = (pd5km - mean(pd5km))/sd(pd5km)
          )

 
    
# Standardize survey covariates 
    # similar to NDVI, there is only one beta across seasons
    # But, detection probability (and the effect of survey effort on it) is not modeled as varying among seasons, 
    # so we don't need to standardize by season
    # column 'effort' = number of days that a site was sampled
    K$effort_std <- (K$effort - mean(K$effort))/sd(K$effort)
    
    K.std <- data.frame(K) %>% 
      transmute(
        k1.std = (season1 - mean(season1))/sd(season1),
        k2.std = (season2 - mean(season2))/sd(season2),
        k3.std = (season3 - mean(season3))/sd(season3)
      ) %>% as.matrix()
```
```{r pivot site data to a longer format}
# data.site

# start with the seasonally-varying NDVI data
data.site.long <-
  data.site %>% 
    select(site,
           ndvi_wd100m, ndvi_ww100m, ndvi_cw100m,
           ndvi_wd1km, ndvi_ww1km, ndvi_cw1km,
           ndvi_wd5km, ndvi_ww5km, ndvi_cw5km) %>%
    pivot_longer(cols = starts_with("ndvi_"), names_to = "season", 
                 names_prefix = "ndvi_", values_to = "ndvi") %>% # %>% pull(ndvi)
    select(site, season, ndvi)

data.site.long$season <- rep(c("1_warmdry", "2_warmwet", "3_coolwet"), 3*50)
data.site.long$scale <- rep(rep(c("100m", "1km", "5km"), each = 3), 50)

data.site.long$imp <- rep((data.site %>% select(site, imp100m, imp1km, imp5km) %>%
    pivot_longer(cols = starts_with("imp"), names_to = "scale",
                 names_prefix = "imp", values_to = "imp") %>% arrange(site, scale) %>% pull(imp)), each = 3)

data.site.long$sdhi <- rep((data.site %>% select(site, sdhi100m, sdhi1km, sdhi5km) %>%
    pivot_longer(cols = starts_with("sdhi"), names_to = "scale",
                 names_prefix = "sdhi", values_to = "sdhi") %>% arrange(site, scale) %>% pull(sdhi)), each = 3)

data.site.long$pd <- rep((data.site %>% select(site, pd100m, pd1km, pd5km) %>%
    pivot_longer(cols = starts_with("pd"), names_to = "scale",
                 names_prefix = "pd", values_to = "pd") %>% arrange(site, scale) %>% pull(pd)), each = 3)



# standardize the variables by season 
  # the model is structured to take these covariates in the form of a matrix, 
  # but we will reformat it as such when we get to fitting the model.
data.site.long <- data.site.long %>% mutate(
  imp_std = (imp - mean(imp, na.rm = TRUE))/sd(imp),
  ndvi_std = (ndvi - mean(ndvi, na.rm = TRUE))/sd(ndvi),
  sdhi_std = (ndvi - mean(ndvi, na.rm = TRUE))/sd(ndvi),
  pd_std = (pd - mean(pd, na.rm = TRUE))/sd(pd),
  .by = c(season, scale)
)

data.site.long$impquad <- data.site.long$imp_std^2

data.site.long
```


```{r import species trait data}
# Species Covariates
        data.spp <- read.csv("~/GitHub/metrophxwildlife-communitydynamics/data/model_inputs/EltonTraits_22mammalbird.csv") %>%
                  arrange(name_short)  # this is the column on which the data is all ordered
        
# create new data columns astandardized versions
    # BodyMass.Value = average adult body mass for each species
    # logmass = log-transformed average body mass
    # pc_carn = carnivory, the percentage of diet consisting of vertebrate prey
    
data.spp$name_short[which(data.spp$Species == "VUMA")] <- "kit_fox" # this species name somehow got an extra space
```
```{r select species and standardize traits}

# Restrict which species to be included the model
  spp.model <- data.spp %>% 
    filter(taxon_class == "Mammalia") %>%   # select only mammal species
    # filter(!name_short %in% c("quail_gambels", "roadrunner")) %>% # drop specific species
    pull(name_short) 


# Subset the species data
  y <- y[,,,which(dimnames(y)[[4]] %in% spp.model)] #%>% str() 
  ysum <- ysum[,,which(dimnames(ysum)[[3]] %in% spp.model)]
  data.spp <- data.spp %>% filter(name_short %in% spp.model)
  n.spp <- dim(y)[4]

# Create columns for standardized species traits
  data.spp <- data.spp %>%   # standardize body mass
    mutate(BodyMass_std = (BodyMass.Value -mean(BodyMass.Value))/sd(BodyMass.Value),
          logmass_std = (logmass - mean(logmass))/sd(logmass),
          carn_std = (pc_carn - mean(pc_carn))/sd(pc_carn),
          dietdiv_std = (dietdiv - mean(dietdiv))/sd(dietdiv)) %>%
    arrange(name_short)   
        
        
# Common names for the species
  names.common <- data.spp$name
  dimnames(y)[[4]] # these are the names by which the occupancy data is alphabetized
  # names.short <- dimnames(y)[[4]]  
  # names.short <- data.spp$name_short     # same thing

# Vectorize phylogenetic structure
  # we need a vector of integers (length = # of species), where each number corresponds to the taxon
  # taxonomic class
  data.spp <- data.spp %>% arrange(taxon_class)
  unique(data.spp$taxon_class)
  t.classes <- unique(data.spp$taxon_class)
  data.spp$class_vec <- as.integer(factor(data.spp$taxon_class, labels = 1:length(t.classes)))  
  
  # taxonomic orders
  data.spp <- data.spp %>% arrange(taxon_order)
  unique(data.spp$taxon_order)
  t.orders <- unique(data.spp$taxon_order)
  data.spp$order_vec <- as.integer(factor(data.spp$taxon_order, labels = 1:length(t.orders)))  

  # taxonomic families
  data.spp <- data.spp %>% arrange(taxon_family)
  unique(data.spp$taxon_family)
  t.families <- unique(data.spp$taxon_family)
  data.spp$family_vec <- as.integer(factor(data.spp$taxon_family, labels = 1:length(t.families)))  


data.spp %>% arrange(dietdiv)

# we don't need the "t." objects anymore
rm(t.classes, t.families, t.orders)

# Look the collinearity of the species traits
cor(data.spp$logmass, data.spp$dietdiv)   # -0.45 isn't too strong
ggplot(data.spp, aes(x = logmass, y = dietdiv)) +
  theme_bw() +
  geom_point()+
  geom_text_repel(aes(label = Species)) +
  labs(x = "log(Body Mass)", y = "Diet Diversity")+
  scale_x_continuous(trans = "log10", breaks = log10(c(0.1, 1, 10)*1000), labels = c("0.1","1", "10"))      
```
```{r summarize naive richness}
# get an array of 1-0-NA detections of each site X season X species
  # ysum01 <- ysum           # approach A
  # ysum01[ysum01 > 1] <- 1
  ysum01 <- apply(y, MARGIN = c(1,3,4), function(x) max(x, na.rm = T)) # approach B
  
  str(ysum01)

# calculate the richness of the species detected in each season and add it to the site covariate dataset
  data.site$rich_obs_wd <- apply(ysum01[,1,], c(1), function(x) sum(x, na.rm = T))
  data.site$rich_obs_ww <- apply(ysum01[,2,], c(1), function(x) sum(x, na.rm = T))
  data.site$rich_obs_cw <- apply(ysum01[,3,], c(1), function(x) sum(x, na.rm = T))
```


# Export Model Input Data
```{r}
# Remove any objects from the environment that aren't necessary to run the models
# e.g.
rm(ysum01)

save.image("~/GitHub/metrophxwildlife-communitydynamics/data/model_inputs/ModelInputData_DCM.RData")
```

# BONUS: Visualize Naive (Observed) Species Richness vs. Environmental Covariates
This is mainly as a preview of the relationships that will be assessed once imperfect detection has been accounted for in the estimates of species richness. Since these are pairwise correlations, the collinearity among the covariates is not being accounted for and these should be taken with a big grain of salt. For instance, vegetation greenness is negatively related to urbanization in the cool-wet season. Certain relationships might not show as substantial (e.g., richness vs. heterogeneity) because they are being overshadowed by the stronger relationships (e.g., richness vs. urbanization).

### Richness vs. Urbanization
Unsurprisingly, richness is negatively correlated with urbanization across all seasons. So far, it does not appear as if there are substantial among-season differences in richness across the urbanization gradient. We'll see if accounting for imperfect detection will change that.
```{r}
ggplot(data = data.site)+
  theme_classic() +
  stat_smooth(aes(x = imp1km, y = rich_obs_wd), fill = "green4", col = "green4",
              method = "lm", formula = y ~ x + I(x^2)) +
  geom_point(aes(x = imp1km, y = rich_obs_wd), col = "green4")+
  stat_smooth(aes(x = imp1km, y = rich_obs_ww), fill = "red4", col = "red4",
              method = "lm", formula = y ~ x + I(x^2)) +
  geom_point(aes(x = imp1km, y = rich_obs_ww), col = "red4")+
  stat_smooth(aes(x = imp1km, y = rich_obs_cw), fill = "blue4", col = "blue4",
              method = "lm", formula = y ~ x + I(x^2)) +
  geom_point(aes(x = imp1km, y = rich_obs_cw), col = "blue4") +
  coord_cartesian(ylim = c(0,12.5))
cor(data.site$imp1km, data.site$rich_obs_wd)
cor(data.site$imp1km, data.site$rich_obs_ww)
cor(data.site$imp1km, data.site$rich_obs_cw)
```

### Richness vs. Landscape Heterogeneity
Generally this shows a weak but positive relationship, with little variation among seasons
```{r}
ggplot(data = data.site)+
  theme_classic() +
  stat_smooth(aes(x = sdhi5km, y = rich_obs_wd), fill = "green4", col = "green4",
              method = "lm") +    # , formula = y ~ x + I(x^2)
  geom_point(aes(x = sdhi5km, y = rich_obs_wd), col = "green4")+
  stat_smooth(aes(x = sdhi5km, y = rich_obs_ww), fill = "red4", col = "red4",
              method = "lm") +
  geom_point(aes(x = sdhi5km, y = rich_obs_ww), col = "red4")+
  stat_smooth(aes(x = sdhi5km, y = rich_obs_cw), fill = "blue4", col = "blue4",
              method = "lm") +
  geom_point(aes(x = sdhi5km, y = rich_obs_cw), col = "blue4") +
  coord_cartesian(ylim = c(0,12.5))
cor(data.site$sdhi5km, data.site$rich_obs_wd)
cor(data.site$sdhi5km, data.site$rich_obs_ww)
cor(data.site$sdhi5km, data.site$rich_obs_cw)
```
### Richness vs. Vegetation Greenness
```{r}
ggplot(data = data.site)+
  theme_classic() +
  stat_smooth(aes(x = ndvi_wd1km, y = rich_obs_wd), fill = "green4", col = "green4",
              method = "lm") +
  geom_point(aes(x = ndvi_wd1km, y = rich_obs_wd), col = "green4")+
  stat_smooth(aes(x = ndvi_ww1km, y = rich_obs_ww), fill = "red4", col = "red4",
              method = "lm") +
  geom_point(aes(x = ndvi_ww1km, y = rich_obs_ww), col = "red4")+
  stat_smooth(aes(x = ndvi_cw1km, y = rich_obs_cw), fill = "blue4", col = "blue4",
              method = "lm") +
  geom_point(aes(x = ndvi_cw1km, y = rich_obs_cw), col = "blue4") +
  coord_cartesian(ylim = c(0,12.5))
cor(data.site$ndvi_wd1km, data.site$rich_obs_wd)
cor(data.site$ndvi_ww1km, data.site$rich_obs_ww)
cor(data.site$ndvi_cw1km, data.site$rich_obs_cw)
```

